---
title: "Chapter 1: Functions in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Writing Functions in R

```{r}
square <- function(x) x**2
```

```{r}
square(1:5)
```

```{r}
rescale <- function(x) {
  m <- mean(x)
  s <- sd(x)
  (x - m) / s
}
```

```{r}
rescale(1:5)
```

## Named Parameters and Default Parameters 

```{r}
rescale <- function(x, only_translate = FALSE) {
  m <- mean(x)
  translated <- x - m
  if (only_translate) return(translated)
  s <- sd(x)
  translated / s
}
```

```{r}
rescale(1:5, only_translate = TRUE)
rescale(1:5, only_translate = FALSE)
```

## The “Gobble Up Everything Else” Parameter: `...`

```{r}
rescale <- function(x, only_translate = FALSE, ...) {
  m <- mean(x, ...)
  translated <- x - m
  if (only_translate) return(translated)
  s <- sd(x, ...)
  translated / s
}
```

```{r}
rescale(c(NA, 1:4))
```

```{r}
rescale(c(NA, 1:4), na.rm = TRUE)
```

```{r}
f <- function(x) x
g <- function(x, ...) x
```

```{r, warning = FALSE}
f(1:4, foo = "bar")
```

```{r}
g(1:4, foo = "bar")
```

```{r}
f <- function(...) list(...)
g <- function(x, y, ...) f(...)
```

```{r}
g(x = 1, y = 2, z = 3,w = 4)
```

```{r}
parameters <- function(...) eval(substitute(alist(...)))
```

```{r}
parameters(a = 4, b = a**2)
```

## Functions Don’t Have Names

```{r}
(function(x) x**2)(4)
```

## Lazy Evaluation

```{r}
f <- function(a, b) a
```

```{r}
f(2, stop("error if evaluated"))
```

```{r}
f(stop("error if evaluated"), 2)
```

You can write default values that depend on other parameters:

```{r}
f <- function(a, b = a) a + b
f(a = 2)
```

```{r}
a <- 4
f <- function(x) {
  a <- 2
  x
}
f(1 + a)
```

```{r}
f <- function(a) function(b) a + b
# f <- function(a) {
#   function(b) {
#     a + b
#   }
# }
```

```{r}
g <- f(2)
g(3)
```

... and a bunch of other stuff

## Vectorized Functions

```{r}
x <- 1:5
y <- 6:10
x - y
```


```{r}
2 * x
```

```{r}
x <- 1:6
y <- 1:3
x - y
```

`compare` function for single values:
```{r}
compare <- function(x, y) {
  if (x < y) {
    -1
  } else if (y < x) {
    1
  } else {
    0
  }
}
compare(1:6, 1:3)
```

`compare` function for vectors:
```{r}
compare <- function(x, y) {
  ifelse(x < y, -1, ifelse(x > y, 1, 0))
}
compare(1:6, 1:3)
```

`compare` using `Vectorize()`:
```{r}
compare <- function(x, y) {
  if (x < y) {
    -1
  } else if (y < x) {
    1
  } else {
    0
  }
}
compare <- Vectorize(compare)
compare(1:6, 1:3)
```

`Vectorize()` vectorizes *_all_* arguments of a function.
```{r}
scale_with <- function(x, y) {
  (x - mean(y)) / sd(y)
}
```

```{r}
scale_with(1:6, 1:3)
```

So running `Vectorize()` on the following function would break it because only `x` should vectorized, not `y`.
```{r}
scale_with <- Vectorize(scale_with)
scale_with(1:6, 1:3)
```

To fix this, use the `vectorize.args` argument:
```{r}
scale_with <- function(x, y) {
  (x - mean(y)) / sd(y)
}
scale_with <- Vectorize(scale_with, vectorize.args = "x")
scale_with(1:6, 1:3)
```

`Vectorize` comes in handy for functions that are complex enough that they're hard to vectorize.

As an example, look at this tree data structure and a function that returns a tree's depth.
```{r}
make_node <- function(name, left = NULL, right = NULL) {
  list(name = name, left = left, right = right)
}

tree <- 
  make_node(name = "root", 
            left = make_node(name = "C", 
                             left = make_node("A"),
                             right = make_node("B")),
            right = make_node(name = "D"))
```

To get tree depth, traverse the tree recursively (prioritizing left node).
```{r}
node_depth <- function(tree, name, depth = 0) {
  if (is.null(tree)) return(NA)
  if (tree$name == name) return(depth)
  
  left <- node_depth(tree$left, name, depth = depth + 1)
  if (!is.na(left)) return(left)
  right <- node_depth(tree$right, name, depth = depth + 1)
  return(right)
}

node_depth(tree, "A")
node_depth(tree, "B")
node_depth(tree, "C")
node_depth(tree, "D")
node_depth(tree, "E")
node_depth(tree, "root")
```

As it is, `node_depth` isn't vectorized.
```{r}
node_depth(tree, c("A", "B", "C", "D"))
```

So, use `Vectorize` to vectorize the `names` argument in `node_depth`.
```{r}
node_depth <- Vectorize(node_depth, 
                        vectorize.args = "name",
                        USE.NAMES = FALSE)
```

```{r}
node_depth(tree, c("A", "B", "C", "D"))
```

## Infix Operators

## Replacement Functions

There is a whole class of such functions you can use to make the appearance of modifying an object, without actually doing it of course. These are called replacement functions and have names that end in `<-`. An example is the `names<-` function.

```{r}
x <- 1:4
x
names(x) <- letters[1:4] # actually the `names<-` function
x
names(x)
```

`names<-` function
```{r}
names(x) <- letters[1:4]
# ... is the same as ...
x <- `names<-`(x, letters[1:4])
```

`attr<-` function
```{r}
x <- 1:4
attributes(x)

attributes(x) <- list(foo = "bar")
attributes(x)

attr(x, "baz") <- "qux"
attributes(x)
```

We can write a replacement function to make the tree construction we had earlier in the chapter slightly more readable. The original was...
```{r}
tree <- make_node("root",
                  make_node("C", make_node("A"),
                                 make_node("B")),
                  make_node("D"))
```

... but we can make functions for setting the child of a node...
```{r}
`left<-` <- function(node, value) {
  node$left = value
  node
}

`right<-` <- function(node, value) {
  node$right = value
  node
}
```

... and then build a tree like this...
```{r}
A <- make_node("A")
B <- make_node("B")
C <- make_node("C")
D <- make_node("D")
root <- make_node("root")
left(C) <- A
right(C) <- B
left(root) <- C
right(root) <- D
tree <- root
```

```{r}
print_tree <- function(tree) {
  build_string <- function(node) {
    if (is.null(node$left) && is.null(node$right)) {
        node$name
    } else {
        left <- build_string(node$left)
        right <- build_string(node$right)
        paste0("(", left, ",", right, ")")
    }
  }
  build_string(tree)
}
print_tree(tree)
```
































